Red-Giant Hot Jupiters: Brilliant Radio Beacons?

### 2015-08-26 ms.tex
incremental updates, not longer just assuming 70 MHz.

%%%%%%%%%% 
%\begin{eqnarray}
%B   &\sim &  10~{\rm G} \left( \frac{t}{4.5~\rm{Gyr}} \right)^{-0.43} \left( \frac{M_p }{M_{p,J}} \right)^{1.38} \label{eq:scalingB} \\
%B   &\sim &  10~{\rm G} \left( \frac{t}{4.5~\rm{Gyr}} \right)^{-0.43} \left( \frac{M_p }{M_{p,J}} \right)^{1.04} \label{eq:scalingB} \\
%f_{\rm cyc} &\sim &  28~{\rm MHz} \left( \frac{t}{4.5~\rm{Gyr}} \right)^{-0.43} \left( \frac{M_p }{M_{p,J}} \right)^{1.38} \label{eq:scalingfc} 
%f_{\rm cyc} &\sim &  28~{\rm MHz} \left( \frac{t}{4.5~\rm{Gyr}} \right)^{-0.43} \left( \frac{M_p }{M_{p,J}} \right)^{1.04} \label{eq:scalingfc} 
%\end{eqnarray}
%%%%%%%%%%
%under this model. 
Note that the cyclotron frequency of Jovian planets typically falls between 10~MHz, the plasma cut-off frequency of Earth's ionosphere, and less than 1~GHz. 
In this regime, there are a number of current and near-future radio wave detectors including UTR-2, LOFAR, LWA, and SKA.

@@

For the rest of the instruments considered, we assume $T_{\rm sys}\approx T_{\rm sky}$,
and use the fit from \cite{Rogers+Bowman2008} to data taken in the 100-200~MHz range:
\begin{equation}
T_{\rm sky} = T_{150} \left( \frac{\nu}{150~\rm MHz}\right)^{-\beta} + T_{\rm CMB} 
\end{equation}
where $T_{150} = 283.2~\rm K$ is the $T_{\rm sky}$ at 150~MHz, $\beta=2.47$,
and $T_{\rm CMB} = 2.73~\rm K$ is the contribution from the cosmic microwave 
background (CMB).
The RMS noise ${\sigma}$ in an interferometric observation, 
integration time $\tau$, bandwidth $\Delta\nu$, effective area $A_{\rm eff}$ 
($\approx0.7 \pi (D/2)^2$ for interferometers comprising antennas with dish size $D$),
and $N$ elements is \citep[see e.g.][]{condon+ransom2016} 
\begin{equation}
\sigma = \frac{2 k T_{\rm sys}}{A_{\rm eff} \, \sqrt{\Delta\nu \, \tau N (N-1)}} 
\left(\frac{\rm 10^{26} Jy}{\rm W m^{-2} Hz^{-1}}\right)
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table*}[htp]
\caption{Detectability of Hot Jupiters with Current and Future Radio Telescopes.}
\begin{center}
\begin{tabular}{c|ccccc} \hline \hline\\[-0.5pc]
Instrument & Band & Sensitivity & $\tau$  & Resolution & $5\sigma_c$\\
 & MHz & $\mu$Jy~hr$^{1/2}$ &  hr & \arcsec\ & $\mu$Jy bm$^{-1}$ \\[0.25pc] 
\hline\\[-0.5pc]
GMRT      & 130-190 & 430   &  460             &  20  & 1070 \\ 
HERA      &  30-80  & 540   &  730             & 880  & $1.4\times10^6$\\ 
LOFAR-LBA &  30-80  & 700   & 1225             & 8.3  & 120 \\ 
LOFAR-HBA & 110-200 &  35   &    3             & 3.3  & 2.6\\  
LWA       &  30-80  & 36000 &  3.2$\times10^6$ & 10.3 & 211\\
MWA       &  30-80  & 18000 &  8.1$\times10^5$ & 1030 & $1.4\times10^6$ \\
SKA-Low   & 50-350 &  3.5   &  0.03            & 7.9  & 57 \\
UTR-2 & & & \\
VLA P-band & 270-470  & 138  & 4760 & \\
\hline
\end{tabular}
\end{center}
\label{tab:sens}
\tablecomments{Time estimates to reach 20~$\mu$Jy RMS.}
\end{table*}

>>>>

%%%%%%%%%% 
%\begin{eqnarray}
%B   &\sim &  10~{\rm G} \left( \frac{t}{4.5~\rm{Gyr}} \right)^{-0.43} \left( \frac{M_p }{M_{p,J}} \right)^{1.38} \label{eq:scalingB} \\
%B   &\sim &  10~{\rm G} \left( \frac{t}{4.5~\rm{Gyr}} \right)^{-0.43} \left( \frac{M_p }{M_{p,J}} \right)^{1.04} \label{eq:scalingB} \\
%f_{\rm cyc} &\sim &  28~{\rm MHz} \left( \frac{t}{4.5~\rm{Gyr}} \right)^{-0.43} \left( \frac{M_p }{M_{p,J}} \right)^{1.38} \label{eq:scalingfc} 
%f_{\rm cyc} &\sim &  28~{\rm MHz} \left( \frac{t}{4.5~\rm{Gyr}} \right)^{-0.43} \left( \frac{M_p }{M_{p,J}} \right)^{1.04} \label{eq:scalingfc} 
%\end{eqnarray}
%%%%%%%%%%
%under this model. 
Note that the cyclotron frequency of Jovian planets typically falls between 10~MHz, the plasma cut-off frequency of Earth's ionosphere, and less than 1~GHz. 
In this regime, there are a number of current and near-future radio wave detectors including GMRT, LOFAR, LWA, MWA, Hera, SKA, and upgrades to the VLA.

@@

The low frequency component of the  Square Kilometer Array (SKA), recently rebaselined 
to have 0.4~km$^2$ collecting area in 130,000 elements with baselines extending
to 65~km, will operate from 50-350~MHz.\footnote{
\url{https://www.skatelescope.org/news/worlds-largest-radio-telescope-near-construction/}}
We estimate the noise in the lower half of the band ($\nu<200~\rm MHz$) assuming
$T_{\rm sys} \approx T_{\rm sky}$, and use the fit from \cite{Rogers+Bowman2008}:
\begin{equation}
T_{\rm sky} = T_{150} \left( \frac{\nu}{150~\rm MHz}\right)^{-\beta} + T_{\rm CMB} 
\end{equation}
where $T_{150} = 283.2~\rm K$ is the $T_{\rm sky}$ at 150~MHz, $\beta=2.47$,
and $T_{\rm CMB} = 2.73~\rm K$ is the contribution from the cosmic microwave 
background (CMB).
While this relation was fit to data taken in the 100-200~MHz range, the authors
found it to be consistent with published measurements from 10-408~MHz.

For the rest of the instruments considered, we also assume $T_{\rm sys} \approx T_{\rm sky}$,
and adopt the \cite{Rogers+Bowman2008} relation.  We caution that is assumption is 
optimistic, and the time estimates could in reality be a factor of a few worse.

The RMS noise ${\sigma}$ in an interferometric observation, 
integration time $\tau$, bandwidth $\Delta\nu$, effective area $A_{\rm eff}$ 
($\approx0.7 \pi (D/2)^2$ for interferometers comprising antennas with dish size $D$),
and $N$ elements is \citep[see e.g.][]{condon+ransom2016} 
\begin{equation}
\sigma = \frac{2 k T_{\rm sys}}{A_{\rm eff} \, \sqrt{\Delta\nu \, \tau N (N-1)}} 
\left(\frac{\rm 10^{26} Jy}{\rm W m^{-2} Hz^{-1}}\right)
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table*}[htp]
\caption{Detectability of Hot Jupiters with Current and Future Radio Telescopes.}
\begin{center}
\begin{tabular}{c|ccccc} \hline \hline\\[-0.5pc]
Instrument & Band & Sensitivity & $\tau$  & Resolution & $5\sigma_c$\\
 & MHz & $\mu$Jy~hr$^{1/2}$ &  hr & \arcsec\ & $\mu$Jy bm$^{-1}$ \\[0.25pc] 
\hline\\[-0.5pc]
GMRT      & 130-190 & 430   &  460             & 16.5 & 280 \\ 
HERA      &  30-80  & 790   &  730             & 1030 & $2.1\times10^6$\\ 
LOFAR-LBA &  30-80  & 700   & 1225             & 8.3  & 211 \\ 
LOFAR-HBA & 110-200 &  35   &    3             & 3.3  & 5.5 \\  
LWA       &  30-80  & 53000 &  3.2$\times10^6$ & 10.3 & 211 \\
MWA       &  30-80  & 27000 &  8.1$\times10^5$ & 1030 & $2.1\times10^6$ \\
SKA-Low   & 50-350  &  4.9   &  0.03            & 7.9  & 57 \\
VLA       & 30-80   & 138  & 4760 & \\
\hline
\end{tabular}
\end{center}
\label{tab:sens}
\tablecomments{Time estimates to reach 20~$\mu$Jy RMS.}
\end{table*}
